---
title: "Danish Kings Homework Week 10"
author: "Amanda, Laura, Cecilie, Sarah"
date: "12/5-2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The task here is to load your Danish Monarchs csv into R using the `tidyverse` toolkit, calculate and explore the kings' duration of reign with pipes `%>%` in `dplyr` and plot it over time.

## Load the kings

Make sure to first create an  `.Rproj` workspace with a `data/` folder where you place either your own dataset or the provided `kings.csv` dataset. 

1. Look at the dataset that are you loading and check what its columns are separated by? (hint: open it in plain text editor to see)

List what is the

separator: semicolon  

2. Create a `kings` object in R with the different functions below and inspect the different outputs.
-  `read.csv()`
-  `read_csv()`
-  `read.csv2()` 
-  `read_csv2()`
```{r eval=TRUE}
# FILL IN THE CODE BELOW and review the outputs

library(tidyverse)

kings1 <- read.csv("~/BA Historie/Digitale metoder/DanishMonarchs-main/data/Kongerækken_clean.csv")
glimpse(kings1)
kings2 <- read_csv("~/BA Historie/Digitale metoder/DanishMonarchs-main/data/Kongerækken_clean.csv")
glimpse(kings2)
kings3 <- read.csv2("~/BA Historie/Digitale metoder/DanishMonarchs-main/data/Kongerækken_clean.csv")
glimpse(kings3)
kings4 <- read_csv2("~/BA Historie/Digitale metoder/DanishMonarchs-main/data/Kongerækken_clean.csv")
glimpse(kings4)
```

Questions: 
1. Which of these functions is a `tidyverse` function? Read data with it below into a `kings` object

Answer: You can tell that kings 2 and 4 are tidyverse functions, as they use "_" instead of "."

2. What is the result of running `class()` on the `kings` object created with a tidyverse function.

Answer: When running class on kings, it shows "spec_tbl_df" "tbl_df" "tbl" "data.frame", which describes a tibble

3. How many columns does the object have when created with these different functions?

Answer: When using function ncol(), there are 6 columns. 

4. Show the dataset so that we can see how R interprets each column

```{r eval=TRUE}
kings <- read_csv2("~/BA Historie/Digitale metoder/DanishMonarchs-main/data/Kongerækken_clean.csv", na="na")

class(kings)


ncol(kings)


glimpse(kings)
```



## Calculate the duration of reign for all the kings in your table
You can calculate the duration of reign in years with `mutate` function by subtracting the equivalents of your `startReign` from `endReign` columns and writing the result to a new column called `duration`. But first you need to check a few things:

* Is your data messy? Fix it before re-importing to R
* Do your start and end of reign columns contain NAs? Choose the right strategy to deal with them: `na.omit()`, `na.rm=TRUE`, `!is.na()`


Create a new column called  `duration` in the kings dataset, utilizing the `mutate()` function from tidyverse. Check with your group to brainstorm the options.

```{r}
# Remove the missing values
kings_duration <- kings %>%
  mutate(duration = reg_slut - reg_start) %>%
  na.omit(duration)

```

## Calculate the average duration of reign for all rulers
Do you remember how to calculate an average on a vector object? If not, review the last two lessons and remember that a column is basically a vector. So you need to subset your `kings` dataset to the `duration` column. If you subset it as a vector you can calculate average on it with `mean()` base-R function. If you subset it as a tibble, you can calculate average on it with `summarize()` tidyverse function. Try both ways!

* You first need to know how to select the relevant `duration` column. What are your options?
* Is your selected `duration` column a tibble or a vector? The `mean()` function can only be run on a vector. The `summarize()` function works on a tibble.
* Are you getting an error that there are characters in your column? Coerce your data to numbers with `as.numeric()`.
* Remember to handle NAs: `mean(X, na.rm=TRUE)`


```{r}
# Finding the mean with the data as a vector object
mean(kings_duration$duration, na.rm=TRUE)

#Answer: The mean is 21

#It didn't work with the summarize function

```

## How many and which kings enjoyed a longer-than-average duration of reign?

You have calculated the average duration above. Use it now to `filter()` the `duration` column in  `kings` dataset. Display the result and also  count the resulting rows with `count()`

```{r}
# Filtering the duration above the mean, which is 21
duration_above_average <- kings_duration %>%
  filter(duration>21)

glimpse(duration_above_average)

```
```{r}
# Counting how many monarchs had a reign duration above average
count(duration_above_average)

# 20 monarchs had a longer than average duration of reign
```


## How many days did the three longest-ruling monarchs rule? 

* Sort kings by reign `duration` in the descending order. Select the three longest-ruling monarchs with the `slice()` function
* Use `mutate()` to create `Days` column where you calculate the total number of days they ruled
* BONUS: consider the transition year (with 366 days) in your calculation!

```{r}
# Sorting kings by reign duration in descending order
duration_descending <- kings_duration %>%
  arrange(desc(duration))

glimpse(duration_descending)

# Selecting the three longest-ruling monarchs 
Longest_rulers <- duration_descending %>%
  slice(c(1,2,3))

glimpse(Longest_rulers)

# Create column with number of days ruled
kings_days <- Longest_rulers %>% 
  mutate(days=duration*365)

glimpse(kings_days)
```


## Challenge: Plot the kings' duration of reign through time

What is the long-term trend in the duration of reign among Danish monarchs? How does it relate to the historical violence trends ? 

* Try to plot the duration of reign column in `ggplot` with `geom_point()` and  `geom_smooth()`
* In order to peg the duration (which is between 1-99) somewhere to the x axis with individual centuries, I recommend creating a new column `midyear` by adding to `startYear` the product of `endYear` minus the `startYear` divided by two (`startYear + (endYear-startYear)/2`).
* Now you can plot the kings dataset, plotting `midyear` along the x axis and `duration` along y axis
* BONUS: add a title, nice axis labels to the plot and make the theme B&W and font bigger to make it nice and legible!

```{r}
# Create midyear column
kings_for_ggplot <- kings_duration %>% 
  mutate(midyear=reg_start+duration/2)

# Create ggplot
kings_for_ggplot %>%
    ggplot(aes(x = midyear, y =duration))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  ggtitle("Kings duration of reign through time")
```

And to submit this rmarkdown, knit it into html. But first, clean up the code chunks, adjust the date, rename the author and change the `eval=FALSE` flag to `eval=TRUE` so your script actually generates an output.
Well done!


