---
title: "Wordclouds of Mette Rosenkrantz"
author: "Laura, Amanda, Sarah, Cecilie"
date: "2025-05-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(tidytext)
library(ggwordcloud)
library(here)
library(pdftools)
library(lubridate)
```


```{r}

# Has been made with help from chatgpt to combine the files to use thhe function "analyze()"
library(pdftools)

analyze <- function(file_path) {
  text <- pdf_text(file_path)
  if (length(text) >= 2) {
    cat(text[2])
  } else {
    cat("PDF has fewer than 2 pages.\n") #only wishing for the second page on each document
  }
}
```


```{r}
#Follows the guide from software carpentry 
list.files(path = "data:breve_rosenkrantz", pattern = "pdf")
```
```{r}
#Follows the guide from software carpentry 
list.files(path = "data:breve_rosenkrantz", pattern = "brev_Rosenkrantz", full.names = TRUE)
```
```{r}
#Follows the guide from software carpentry 
filenames <- list.files(path = "data:breve_rosenkrantz",  
                        pattern = "brev_Rosenkrantz-[0-9]{2}.pdf",
                        full.names = TRUE)
for (f in filenames) {
  print(f)
  analyze(f)
}
```

```{r}
#Combining with tidytext
analyze <- function(file_path) {
  text <- pdf_text(file_path)
  
  if (length(text) < 2) {
    return(tibble(file = basename(file_path), word = NA))
  }

  second_page <- text[2]
  #For making it so it only uses the second page
  
  tibble(file = basename(file_path), text = second_page) %>%
    unnest_tokens(word, text)
}
```


```{r}
filenames <- list.files(path = "data:breve_rosenkrantz",  
                        pattern = "brev_Rosenkrantz-[0-9]{2}.pdf",
                        full.names = TRUE)

# Use lapply to process all files and bind results
all_words <- lapply(filenames, analyze) %>%
  bind_rows()
```


```{r removes stopwords}
# Made with help from chatgpt to have our stoplist put in 

my_stops <- readLines("stoplist.txt") %>%
  str_trim() %>%
  str_to_lower()

my_stops_df <- tibble(word = my_stops)

# Normalize word column and remove stopwords
all_words_clean <- all_words %>%
  filter(!is.na(word)) %>%
  mutate(word = str_to_lower(str_trim(word))) %>%
  filter(!str_detect(word, "[[:punct:]]")) %>%
  anti_join(my_stops_df, by = "word")

# Remove numeric words
all_words_clean_no_numeric <- all_words_clean %>% 
  filter(!str_detect(word, "^\\d+$"))


```


Wordcloud for Rosenkrantz.
The following codes is made by Max Odsbjerg Pedersen, but we have modified it for our purposes. 
```{r Count words}

all_words_clean_no_numeric %>% 
  count(word) %>% 
  arrange(-n)


```

Frequence analysis with wordcloud
```{r}

all_words_clean_no_numeric %>% 
  count(word, file, sort=TRUE)

```


Before we calculate the frequence of the words, we need R to count how many words there are in each letter. This is done by using the function `group_by` followed by `summarise`:
```{r}

all_words_clean_no_numeric %>% 
  count(word,file) %>% 
  group_by(file) %>% 
  summarise(total=sum(n))->total_words

```

Then we add the total number of words to our dataframe, which we do with `left_join`:
```{r}

all_words_clean_no_numeric %>% 
  count(word, file, sort=TRUE) %>% 
  left_join(total_words, by="file") ->all_words_clean_no_numeric_count

```


```{r}

all_words_clean_no_numeric_count %>% 
  bind_tf_idf(word,file,n)->all_words_tf_idf

```

We do not see any interesting words, because R lists all the words in an ascending order. Instead, we will ask it to list them in a descending order – highest to lowest tf_idf:
```{r}
all_words_tf_idf %>% 
  arrange(desc(tf_idf))

```

```{r}
# Has been made with help from chatgpt, to help us making the pdf title into titles of dates

file_years <- c(
  "brev_Rosenkrantz-01.pdf" = "Letter 1: 1571-03-17",
  "brev_Rosenkrantz-02.pdf" = "Letter 2: 1571-10-26",
  "brev_Rosenkrantz-03.pdf" = "Letter 3: 1573-02-19",
  "brev_Rosenkrantz-04.pdf" = "Letter 4: 1573-02-22",
  "brev_Rosenkrantz-05.pdf" = "Letter 5: 1578-09-03",
  "brev_Rosenkrantz-06.pdf" = "Letter 6: 1573-03-18",
  "brev_Rosenkrantz-07.pdf" = "Letter 7: 1573-04-18",
  "brev_Rosenkrantz-08.pdf" = "Letter 8: 1573-06-27",
  "brev_Rosenkrantz-09.pdf" = "Letter 9: 1574-01-18"
)

all_words_tf_idf <- all_words_tf_idf %>%
  mutate(year = file_years[file])


```

visualizastion of the wordclouds. Creating a better layout and adding the new names by grouping it as "year". 
```{r}
all_words_tf_idf %>%
  arrange(desc(tf_idf)) %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>%
  group_by(year) %>%
  slice_max(tf_idf, n = 15, with_ties = FALSE) %>%
  ggplot(aes(label = word, size = tf_idf, color = tf_idf)) +
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 10) +
  theme_minimal() +
  facet_wrap(~year, ncol = 4, scales = "free") +
  scale_color_gradient(low = "blue", high = "black") +
  labs(
    title = "Rosenkrantz letters: most unique words per year",
    subtitle = "Importance determined by term frequency (tf) - inversed document frequency (idf)",
    caption = "Data from Det Kongelige Bibliotek"
  )

  
```


Saving the wordcloud as a png
```{r}

ggsave("wordcloudMetteRosenkrantz.png", width = 20, height = 10,  units = "cm", bg = "white")

```










